Maenpuen et al. Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2022-05-03 ms/LMA_method.Rmd
update ms

*Effects of leaf sizes and thicknesses on the relationship between whole-leaf and leaf disc LMA*

To quantify the effect of leaf area and thickness on the relationship between whole-leaf and leaf disc LMA, we built a hierarchical Bayesian model.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of  leaf area and leaf thickness on the log-scale;

\begin{align}
\mathrm{ln}(LMA_{Wi}) & \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i) (\#eq:LMA)\\
\tilde{\mu_i} & = \beta_0 + \beta_1 \mathrm{ln}(LA_{i}) + \beta_2 \mathrm{ln}(LT_{i}) (\#eq:mu)
\end{align}

where $LMA_{Wi}$, $LMA_{Di}$, $LA_{i}$, $LT_{i}$ are the whole-leaf LMA, leaf disc LMA, leaf area and leaf thickness for tree *i*, respectively.
$\tilde{\mu_i}$ is the effect of leaf area and leaf thickness on the mean estimates for the whole-leaf LMA of tree *i*, and $\sigma_i$ is the variance of whole-leaf LMA for tree *i*, and $\beta$ are coefficients.

The variance of whole-leaf LMA ($\sigma_i$) was assumed to follow a normal distribution;

\begin{align}
\mathrm{ln}(\sigma_i) & \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega) (\#eq:ln-sigma) \\
\tilde{\sigma_i} & = \gamma_0 + \gamma_1 \mathrm{ln}(LA_{i}) + \gamma_2 \mathrm{ln}(LT_{i}) (\#eq:tilde-sigma)
\end{align}

where $\tilde{\sigma_i}$ is the effect of leaf area and leaf thickness on the variance of whole-leaf LMA,
$\omega$ is a scaling hyperparameter, and $\gamma$ are coefficients.
To allow comparisons among parameter estimates, $\mathrm{ln}(LA)$ and $\mathrm{ln}(LT)$ were scaled to a mean of 0 and standard deviation of 1.

@@

Bivariate trait relationships between disc-based and whole-leaf log-transformed LMA were summarized with variance explained (*R^2^*) based on Pearson correlations and with standardized major axis (SMA) regressions [@Warton2006] across all the samples and different leaf area and thickness groups.
Species and individual groups were divided into four categories based on the medians of leaf area and leaf thickness across all species and all individuals, respectively:
thin and small-leaved, thin and large-leaved, thick and small-leaved, thick and large-leaved species (individuals).
Median values are as following:
small-leaved species (LA < `r val$LA_mid` cm^2^),
thin-leaved species (LT < `r val$LT_mid` mm),
small-leaved individuals (LA < `r val$LA_mid2` cm^2^),
and thin-leaved individuals (LT < `r val$LT_mid2` mm).

@@

Disc-based estimates of species mean LMA matched well whole-leaf estimates (Table 1; *R^2^* = `r val$r2_LMA` for all species).
Species with thicker and smaller leaves had stronger correlations between disc-based and whole-leaf estimates of LMA (Table 1, Fig. 1).
There was a negative correlation between species mean leaf area and leaf thickness (*r* = `r val$LA_LT_r`, *P* `r val$LA_LT_p`, *n* = `r val$LA_LT_n`).
When we pooled all the species, there were significant differences in intercepts for species mean LMA (Table 1), suggesting that the whole-leaf estimates of LMA were 9.4% greater than disc-based estimates in the arithmetic scale.
Large-leaved species also had significant positive intercepts, suggesting that whole-leaf estimates of LMA were 22.1% greater than disc-based estimates in the arithmetic scale (Table 1).
Intercepts did not differ for small-leaved species (Table 1).

Disc-based estimates of individual mean LMA also showed positive correlations with whole-leaf estimates but their *R^2^* values were smaller than those for the species means (Fig. 2).
Similar to the species-level pattern, individuals with thicker and smaller leaves had stronger correlations (Fig. 2).
The relationships of the whole-leaf LMA : disc leaf LMA ratio with total dry mass for the leaf disc (we weighted 2-3 discs together), with leaf thickness, and with leaf area were heteroscedastic (unequal variance, *P* < 0.001; Appendix S3).
Samples with small total dry mass, thin leaves, and larger leaves tended to show greater variance (Fig. 2; Appendix S3).
The total dry mass of leaf disc was positively correlated with both leaf thickness and leaf punch size (Appendix S4).
Amounts of intraspecific variation (CV; coefficient of variation) only weakly matched between whole-leaf and disc-based estimates (Fig. 3; *R^2^* = `r val$cv_r2`).

@@

**Fig. 1.**
Relationships between species mean leaf mass per area (LMA) determined by using whole leaves and leaf discs.
Species groups are divided into four categories on the basis of the medians of the leaf size and leaf thickness across all the species.
Dashed lines indicate 1:1 lines.
Blue solid lines indicate standardized major axis regressions.
The 95% confidence intervals are presented as the shaded area.
All the correlations are significant (*P* < 0.001).

![](../figs/lalt_pool_grid.png)

@@

**Fig. 2.**
Relationships between individual leaf mass per area (LMA) determined by using whole leaves and leaf discs.
Individual groups are divided into four categories based on the basis of the medians of the leaf size and leaf thickness across all the individuals.
All the correlations are significant (*P* < 0.001).
Details as in Fig. 1.

![](../figs/lalt_tree_grid.png)

>>>>

*Effects of leaf morphology on the relationship between whole-leaf and leaf disc LMA*

To quantify the effect of leaf morphology on the relationship between whole-leaf and leaf disc LMA, we built a hierarchical Bayesian model.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of  leaf tissue density, leaf area and leaf thickness on the log-scale;

\begin{align}
\mathrm{ln}(LMA_{Wi}) & \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i^2) (\#eq:LMA)\\
\tilde{\mu_i} & = \beta_0 + \beta_1 \mathrm{ln}(LD_{i}) + \beta_2 \mathrm{ln}(LA_{i}) + \beta_3 \mathrm{ln}(LT_{i}) (\#eq:mu)
\end{align}

where $LMA_{Wi}$, $LMA_{Di}$, $LD_{i}$, $LA_{i}$, $LT_{i}$ are the observed whole-leaf LMA, leaf disc LMA, leaf tissue density, leaf area and leaf thickness for tree *i*, respectively,
$\tilde{\mu_i}$ is the effect of leaf tissue density, leaf area and leaf thickness on the mean estimates for the whole-leaf LMA of tree *i*, and $\sigma_i^2$ is the variance of whole-leaf LMA for tree *i*, and $\beta$ are coefficients.
The standard deviation of whole-leaf LMA ($\sigma_i$) was assumed to follow a normal distribution;

\begin{align}
\mathrm{ln}(\sigma_i) & \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega^2) (\#eq:ln-sigma) \\
\tilde{\sigma_i} & = \gamma_0 + \gamma_1 \mathrm{ln}(LD_{i}) + \gamma_2 \mathrm{ln}(LA_{i}) + \gamma_3 \mathrm{ln}(LT_{i}) (\#eq:tilde-sigma)
\end{align}

where $\tilde{\sigma_i}$ is the effect of leaf area and leaf thickness on the variance of whole-leaf LMA,
$\omega^2$ is a scaling hyperparameter, and $\gamma$ are coefficients.
To allow comparisons among parameter estimates, $\mathrm{ln}(LD)$, $\mathrm{ln}(LA)$ and $\mathrm{ln}(LT)$ were scaled to a mean of 0 and standard deviation of 1.
Note that this model does not include measurement errors of leaf disc LMA implemented in the following analysis, because there are too many scaling parameters

@@

Bivariate trait relationships between disc-based and whole-leaf log-transformed LMA were further summarized with variance explained (*R^2^*) based on Pearson correlations and with standardized major axis (SMA) regressions [@Warton2006] across all the samples and different leaf area and thickness groups.

Species groups were divided into four categories based on the medians of leaf area and leaf thickness across all species:
thin and small-leaved, thin and large-leaved, thick and small-leaved, thick and large-leaved species.
Median values are as following:
small-leaved species (LA < `r val$LA_mid` cm^2^),
and thin-leaved species (LT < `r val$LT_mid` mm).

@@

Leaf tissue density, leaf area and leaf thickness influenced the relationship between the whole-leaf LMA and leaf disc LMA (Fig. 1).
Specifically, species with greater leaf tissue density (Fig. 1a, $\beta_1$), and larger (Fig. 1a, $\beta_2$) and thicker leaves (Fig. 1a, $\beta_3$) had greater whole-leaf LMA than leaf disc LMA.
Species with lower leaf tissue density (Fig. 1b, $\gamma_1$), larger leaf area (Fig. 1b, $\gamma_2$), and thinner leaves (Fig. 1b, $\gamma_3$) had larger variance in the relationship between the whole-leaf LMA and leaf disc LMA.

When we pooled all the species, there were significant differences in intercepts for species mean LMA (Table 1), suggesting that the whole-leaf estimates of LMA were 9.4% greater than disc-based estimates in the arithmetic scale.
Large-leaved species also had significant positive intercepts, suggesting that whole-leaf estimates of LMA were 22.1% greater than disc-based estimates in the arithmetic scale (Table 1).
Intercepts did not differ for small-leaved species (Table 1).

Species with less dense, thicker and smaller leaves had stronger correlations between disc-based and whole-leaf estimates of LMA (Table 1, Fig. 1).

There was a negative correlation between species mean leaf area and leaf thickness (*r* = `r val$LA_LT_r`, *P* `r val$LA_LT_p`, *n* = `r val$LA_LT_n`).

The relationships of the whole-leaf LMA : disc leaf LMA ratio with total dry mass for the leaf disc (we weighted 2-3 discs together), with leaf thickness, and with leaf area were heteroscedastic (unequal variance, *P* < 0.001; Appendix S3).
Samples with small total dry mass, thin leaves, and larger leaves tended to show greater variance (Fig. 2; Appendix S3).
The total dry mass of leaf disc was positively correlated with both leaf thickness and leaf punch size (Appendix S4).
Amounts of intraspecific variation (CV; coefficient of variation) only weakly matched between whole-leaf and disc-based estimates (Fig. 3; *R^2^* = `r val$cv_r2`).

@@

**Fig. 1.**
Standardized regression coefficients modelig
(a) the effects of leaf tissue density (LD; $\beta_1$),
leaf area (LA; $\beta_2$),
leaf thickness (LT; $\beta_3$)
on the mean estimates of whole-leaf LMA, and
(b) the effects of leaf tissue density ($\gamma_1$),
leaf area ($\gamma_2$), and
leaf thickness ($\gamma_3$)
on the estimated variance of whole-leaf LMA.
The leaf disc LMA was used for the baseline mean of the whole-leaf LMA.
Circles show posterior means of coefficients.
Thick and thin lines indicate 90% and 95% credible intervals, respectively.

![](../figs/coef_sp3.png)

@@

**Fig. 2.**
Relationships between species mean leaf mass per area (LMA) determined by using whole leaves and leaf discs for (a) nondense-leave species and (b) dense-leaved species.
Species groups are divided into eight categories on the basis of the medians of the leaf tissue density, leaf area and leaf thickness across all the species.
Dashed lines indicate 1:1 lines.
Blue solid lines indicate standardized major axis regressions.
The 95% confidence intervals are presented as the shaded area.
All the correlations are significant (*P* < 0.001).

![](../figs/lalt_sp_grid_dense.png)

### 2022-05-03 ms/LMA_method.tex
update ms

\emph{Effects of leaf sizes and thicknesses on the relationship between whole-leaf and leaf disc LMA}

To quantify the effect of leaf sizes and thicknesses on the relationship between whole-leaf and leaf disc LMA, we built hierarchical Bayes models.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of leaf size and leaf thickness on the log-scale;

\begin{align}
\mathrm{ln}(LMA_{Wi}) & \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i) \label{eq:LMA}\\
\tilde{\mu_i} & = \beta_0 + \beta_1 \mathrm{ln}(LA_{i}) + \beta_2 \mathrm{ln}(LT_{i}) \label{eq:mu}
\end{align}

where \(LMA_{Wi}\), \(LMA_{Di}\), \(LA_{i}\), \(LT_{i}\) are the whole-leaf LMA, leaf disc LMA, leaf area and leaf thickness for tree \emph{i}, respectively.
\(\tilde{\mu_i}\) is the effect of leaf area and leaf thickness on the mean estimates for the whole-leaf LMA of tree \emph{i}, and \(\sigma_i\) is the variance of whole-leaf LMA for tree \emph{i}, and \(\beta\) are coefficients.

The variance of whole-leaf LMA (\(\sigma_i\)) was assumed to follow a normal distribution;

\begin{align}
\mathrm{ln}(\sigma_i) & \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega) \label{eq:ln-sigma} \\
\tilde{\sigma_i} & = \gamma_0 + \gamma_1 \mathrm{ln}(LA_{i}) + \gamma_2 \mathrm{ln}(LT_{i}) \label{eq:tilde-sigma}
\end{align}

where \(\tilde{\sigma_i}\) is the effect of leaf area and leaf thickness on the variance of whole-leaf LMA,
\(\omega\) is a scaling hyperparameter, and \(\gamma\) are coefficients.
To allow comparisons among parameter estimates, \(\mathrm{ln}(LA)\) and \(\mathrm{ln}(LT)\) were scaled to a mean of 0 and standard deviation of 1.

@@

\emph{Bivariate trait relationships}

@@

\emph{Intraspecifc variation}

>>>>

\emph{Effects of leaf morphology on the relationship between whole-leaf and leaf disc LMA}

To quantify the effect of leaf morphology on the relationship between whole-leaf and leaf disc LMA, we built a hierarchical Bayesian model.
The whole-leaf LMA was assumed to follow a normal distribution (N) with the mean of leaf disc LMA and covariates of leaf tissue density, leaf area and leaf thickness on the log-scale;

\begin{align}
\mathrm{ln}(LMA_{Wi}) & \sim N(\mathrm{ln}(LMA_{Di}) + \tilde{\mu_i}, \sigma_i^2) \label{eq:LMA}\\
\tilde{\mu_i} & = \beta_0 + \beta_1 \mathrm{ln}(LD_{i}) + \beta_2 \mathrm{ln}(LA_{i}) + \beta_3 \mathrm{ln}(LT_{i}) \label{eq:mu}
\end{align}

where \(LMA_{Wi}\), \(LD_{i}\), \(LA_{i}\), \(LT_{i}\) are the observed whole-leaf LMA, leaf tissue density, leaf area and leaf thickness for tree \emph{i}, respectively,
\(LMA_{Di}\) is the true unobserved value of leaf disc LMA for tree \emph{i}, which can be modeled using the observed leaf disc LMA.
\(\tilde{\mu_i}\) is the effect of leaf tissue density, leaf area and leaf thickness on the mean estimates for the whole-leaf LMA of tree \emph{i}, and \(\sigma_i^2\) is the variance of whole-leaf LMA for tree \emph{i}, and \(\beta\) are coefficients.

The standard deviation of whole-leaf LMA (\(\sigma_i\)) was assumed to follow a normal distribution;

\begin{align}
\mathrm{ln}(\sigma_i) & \sim N(\mathrm{ln}(\tilde{\sigma_i}), \omega^2) \label{eq:ln-sigma} \\
\tilde{\sigma_i} & = \gamma_0 + \gamma_1 \mathrm{ln}(LD_{i}) + \gamma_2 \mathrm{ln}(LA_{i}) + \gamma_3 \mathrm{ln}(LT_{i}) \label{eq:tilde-sigma}
\end{align}

where \(\tilde{\sigma_i}\) is the effect of leaf area and leaf thickness on the variance of whole-leaf LMA,
\(\omega^2\) is a scaling hyperparameter, and \(\gamma\) are coefficients.
To allow comparisons among parameter estimates, \(\mathrm{ln}(LD)\), \(\mathrm{ln}(LA)\) and \(\mathrm{ln}(LT)\) were scaled to a mean of 0 and standard deviation of 1.

@@

\emph{Bivariate trait relationships for mean and CV}

@@
